---
title: 'Assignment #2'
author: "Cooper Nederhood & Xiangyu Ma"
date: "November 2, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  include = FALSE,
  options(digits = 3))
```

```{r}
library(tidyverse)
library(knitr)
library(MASS)
library(sem)
library(dplyr)
set.seed(47)
```


## Set up data
```{r}
setwd("/home/cooper/Documents/autumn_2018/applied_regression/homeworks/HW_2")
df <- read_csv("usa_00001.csv")
head(df)
print(2+2)
```

## Filter and save data out
```{r}
df_subset <- filter(df, AGE<=40, AGE>=20, SEX==2, STATEICP==71)
write_csv(df_subset, "filtered_women.csv")
```

## Work with filtered data and generate the flags
```{r}
#rm(df)
df_subset <- mutate(df_subset, xy_race = if_else(RACE == 1, "White", 
                                   if_else(RACE == 2, "Black",
                                           if_else(RACE >3 & RACE < 7, "Asian", "Other")))) 
df_subset <- mutate(df_subset, xy_hisp = if_else(HISPAN == 0, 0, 1))

df_subset <- mutate(df_subset, xy_met = if_else(MET2013 == 0, 0, 1))

# recode a bunch of dummy variables for education
df_subset <- mutate(df_subset, less_hs = if_else(EDUCD < 30, 1, 0)) %>%
  mutate(some_hs = if_else(EDUCD >= 30 & EDUCD < 60, 1, 0)) %>%
  mutate(ged = if_else(EDUCD == 64, 1, 0)) %>%
  mutate(hs_grad = if_else(EDUCD == 63, 1, 0)) %>%
  mutate(some_col = if_else((EDUCD > 64 & EDUCD < 81), 1, 
    if_else(EDUCD == 90, 1, 
        if_else(EDUCD == 100, 1, 
            if_else(EDUCD >= 110 & EDUC < 114, 1, 0))))) %>%
  mutate(assoc_deg = if_else(EDUCD > 80 & EDUC <90, 1, 0)) %>%
  mutate(bach = if_else(EDUCD == 101, 1, 0)) %>%
  mutate(masters = if_else(EDUCD == 114, 1, 0)) %>%
  mutate(profd = if_else(EDUCD == 115, 1, 0)) %>%
  mutate(phd = if_else(EDUCD == 116, 1, 0))
```


# Question 1: Run an OLS equation using whether the woman has a child or not against her age, education, race, Hispanic status, and place of residence. Recover the predicted values. Are they within the unit interval?
```{r}
df <- df_subset
names(df)
```


```{r}
# Add flag for having a child or not
df <- mutate(df, has_child = if_else(NCHILD==0, 0, 1))
```

```{r}
lm1 <- lm(has_child ~ AGE + less_hs + some_hs + ged + hs_grad + some_col + assoc_deg + bach + masters + profd + phd + xy_race + xy_hisp + MET2013, df, weights=PERWT)
summary(lm1)
```

```{r}
q1_predicted <- as_tibble(list(pred = lm1$fitted.values))
```


```{r}
ggplot(q1_predicted, aes(x = pred))+
  geom_histogram(binwidth = 0.01)
```

# Question 2: Run the same equation using a logit model rather than the OLS. Recover the predicted values. What is the correlation between the predicted values in question one and two?

```{r}
lm2 <- glm(has_child ~ AGE + less_hs + some_hs + ged + hs_grad + some_col + assoc_deg + bach + masters + profd + phd + xy_race + xy_hisp + MET2013,
    family = binomial("logit"),
    data = df)
summary(lm2)

```






